---
title: "Preprocess validation data"
output: html_notebook
---

Preprocessing external validation data, completely separately from the training and testing data.

# First, we pre-process the external Chopra data.
Chopra data from https://osf.io/tw4ye/ (and paper is here: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5354921/)


```{r}
# load in Chopra data
load('Chopra_external/beta.testing.rda')
beta.testing <- as.data.frame(beta.testing) # hmm this is 282 samples not 272 like the paper says...

# need to get probe positions from a harmonised 450k annotation
# using a random TCGA download that includes the probe locations
library(data.table)
probe_location_data <- fread('jhu-usc.edu_BLCA.HumanMethylation450.24.lvl-3.TCGA-XF-AAMW-01A-11D-A42F-05.gdc_hg38.txt')
probe_location_data <- probe_location_data[, c('Composite Element REF', 'Chromosome', 'Start', 'End')]

beta.testing$probe = row.names(beta.testing)
chopera_validation <- merge(probe_location_data, beta.testing, by.x = 'Composite Element REF', by.y = 'probe', all = TRUE, sort = FALSE)

chopera_GRanges <- GRanges(seqnames =chopera_validation$Chromosome, strand ='*', ranges = IRanges(start = chopera_validation$Start, end = chopera_validation$End))
mcols(chopera_GRanges) <- chopera_validation[, 5:ncol(chopera_validation)] # all the samples

chopera_GRanges[1:5, 1:5]
```

Now we have chopra probe positions, we need the positions of features that the model takes in as input
```{r}
# we saved the positions of the processed training data. lets read them in here

seqnames_and_start <- fread('training_testing_maxgap_50_newdata_processed_m_vals.csv', select = c('seqnames', 'start'), sep = ',')

input_features_Gr <- GRanges(seqnames = seqnames_and_start$seqnames, ranges = IRanges(start = seqnames_and_start$start, end = seqnames_and_start$start), strand = '*')
input_features_Gr
seqlevelsStyle(input_features_Gr) <- seqlevelsStyle(chopera_GRanges)
```

Now overlap the input features to get the right features for Chopra data
```{r}

overlaps <- findOverlaps(chopera_GRanges, input_features_Gr)
overlaps


mean(1:length(input_features_Gr) %in% subjectHits(overlaps)) # good so we overlap for every input_feature
subjectHits(overlaps)[duplicated(subjectHits(overlaps))] # but multiple chopera positions can map to one input_feature position
# lets just delete one, it should be very similar methylation as it is the same position


# need to order because for some reason becomes unordered
ordered_overlaps <- arrange(as.data.frame(overlaps), subjectHits)
ordered_overlaps

chopra_input_Gr <- chopera_GRanges[ordered_overlaps$queryHits][!duplicated(granges(chopera_GRanges[ordered_overlaps$queryHits]))]

# check features are correct
mean(seqnames(chopra_input_Gr) == paste0('chr', seqnames_and_start$seqnames))
mean(start(chopra_input_Gr) == seqnames_and_start$start) # yay both are 1

chopra_input_Gr[1:10,1:10]
```



And also load in testing diagnoses and process
```{r}
load('Chopra_external/pdata.testing.rda')

mean(row.names(pdata.testing) %in% colnames(mcols(chopra_input_Gr)))
colnames(mcols(chopra_input_Gr))[!(colnames(mcols(chopra_input_Gr)) %in% row.names(pdata.testing))]

# need to throw away samples we don't have labelled data for
chopra_input_Gr <- chopra_input_Gr[, colnames(mcols(chopra_input_Gr)) %in% row.names(pdata.testing)]

pdata.testing
chopera_data_info <- data.frame(label = pdata.testing$TissueType, id =  row.names(pdata.testing))

# also want to remove AMLs
to_remove <- chopera_data_info[chopera_data_info$label == 'AML', 'id']
chopra_input_Gr <- chopra_input_Gr[, !(colnames(mcols(chopra_input_Gr)) %in% to_remove)]
chopera_data_info <- chopera_data_info[chopera_data_info$label != 'AML',]

# and remove Other samples
to_remove <- chopera_data_info[chopera_data_info$label == 'Other', 'id']
chopra_input_Gr <- chopra_input_Gr[, !(colnames(mcols(chopra_input_Gr)) %in% to_remove)]
chopera_data_info <- chopera_data_info[chopera_data_info$label != 'Other',]


mean(colnames(mcols(chopra_input_Gr)) == chopera_data_info$id) # wow it is in the same order

chopra_input_Gr[1:10,1:10]
dim(as.data.frame(chopra_input_Gr))
chopera_data_info
```


And now we can save
```{r}

chopra_input_df <- as.data.frame(chopra_input_Gr)
chopra_input_df

mean(chopra_input_df$start == seqnames_and_start$start) # double check rows are in correct order
mean(colnames(chopra_input_df)[6:length(colnames(chopra_input_df))] == paste0('X', chopera_data_info$id)) # double check columns are in correct order


write.csv(chopra_input_df, 'Chopra_external/preprocessed_validation_beta.testing.csv')
write.csv(chopera_data_info, 'Chopra_external/preprocessed_validation_pdata.testing_diagnoses.csv')
```



Now checking the pre-processed validation data - does it look like we pre-processed it right?
```{r}
chopra_input_df <- fread('Chopra_external/preprocessed_validation_beta.testing.csv')
chopera_data_info <- fread('Chopra_external/preprocessed_validation_pdata.testing_diagnoses.csv', header = TRUE)

chopra_input_df[1:5,1:10]

table(chopera_data_info$label)


# comparing the probes to some training probes to see if they look vaguely similar
train <- fread('training_testing_maxgap_50_newdata_processed_m_vals.csv')
train <- train[, 4:ncol(train)]
train_diag <- fread('training_testing_diagnoses_maxgap_50_newdata.csv')
train_diag$overallpath[1:10] # first are 3s = oncocytoma
train[1:5, 1:10] #rows are features (probes)

# oncocytomas in test: 8784241087_R01C01, 8795194005_R02C02, 8795194005_R03C02, 8795194005_R06C01
betas <- chopra_input_df[1:10, c('X8784241087_R01C01', 'X8795194005_R02C02', 'X8795194005_R03C02', 'X8795194005_R06C01')]

log2(betas/(1-betas)) # yes looks similar between samples!



# looking at the PCA to see if there are the normal clusters
no_nas <- na.omit(chopra_input_df[, 7:ncol(chopra_input_df)])

pca_res <- prcomp(t(no_nas))

pca_data <- as.data.frame(pca_res$x)
library(ggplot2)
ggplot(data = pca_data) +
  geom_point(aes(x = PC1, y = PC2, colour = chopera_data_info$label)) 
# looks good, have some clusters

```

This looks like we processed it right.




# Now, pre-processing Stanford external data.
Note that Stanford refers to the Brennan dataset.
```{r}
stanford_data <- readRDS('Brennandata/kevinmethylationdata.rds') 
# note: this data was kindly given to us directly from the authors. You can find this dataset at: https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE156932. This is in a slightly different format, so the preprocessing steps may be slightly different.
stanford_data <- data.frame(stanford_data)

View(head(stanford_data))

# need to get probe positions from a harmonised 450k annotation
library(data.table)
probe_location_data <- fread('jhu-usc.edu_BLCA.HumanMethylation450.24.lvl-3.TCGA-XF-AAMW-01A-11D-A42F-05.gdc_hg38.txt', data.table = F)
probe_location_data <- probe_location_data[, c('Composite Element REF', 'Chromosome', 'Start', 'End')]

stanford_data$probe = row.names(stanford_data)
stanford_data_probe_locs <- merge(probe_location_data, stanford_data, by.x = 'Composite Element REF', by.y = 'probe', all = TRUE, sort = FALSE)

stanford_GRanges <- GRanges(seqnames =stanford_data_probe_locs$Chromosome, strand ='*', ranges = IRanges(start = stanford_data_probe_locs$Start, end = stanford_data_probe_locs$End))
mcols(stanford_GRanges) <- stanford_data_probe_locs[, 5:ncol(stanford_data_probe_locs)] # all the samples


stanford_GRanges[1:5, 1:5]


```


Now we have stanford probe positions, we need the positions of features that the model takes in as input
```{r}
# we saved the positions of the processed training data. lets read them in here
seqnames_and_start <- fread('training_testing_maxgap_50_newdata_processed_m_vals.csv', select = c('seqnames', 'start'), sep = ',')

input_features_Gr <- GRanges(seqnames = seqnames_and_start$seqnames, ranges = IRanges(start = seqnames_and_start$start, end = seqnames_and_start$start), strand = '*')
input_features_Gr
seqlevelsStyle(input_features_Gr) <- seqlevelsStyle(stanford_GRanges)
```

Now overlap the input features to get the right features for Stanford data
```{r}
overlaps <- findOverlaps(stanford_GRanges, input_features_Gr)
overlaps

mean(1:length(input_features_Gr) %in% subjectHits(overlaps)) # good so we overlap for every input_feature
subjectHits(overlaps)[duplicated(subjectHits(overlaps))] # but multiple positions can map to one input_feature position
# lets just delete one, it should be very similar methylation as it is the same position

# need to order because for some reason becomes unordered
ordered_overlaps <- arrange(as.data.frame(overlaps), subjectHits)
ordered_overlaps

stanford_input_Gr <- stanford_GRanges[ordered_overlaps$queryHits][!duplicated(granges(stanford_GRanges[ordered_overlaps$queryHits]))]

# check features are correct
mean(seqnames(stanford_input_Gr) == paste0('chr', seqnames_and_start$seqnames))
mean(start(stanford_input_Gr) == seqnames_and_start$start) # yay both are 1

stanford_input_Gr[1:5,1:5]

for (c in colnames(mcols(stanford_input_Gr))) { # some features are always NA (189), this suggests they couldn't be mapped to input features. This also shows some samples have a very high NA number, the worst is X200394970099_R06C01 with 3716 NAs.
  print(sum(is.na(mcols(stanford_input_Gr)[, c])))
}
```

Now get the diagnoses and process
```{r}
stanford_diagnoses <- read.table('Brennandata/Kevinclinicaldata.csv', sep = ',', header = TRUE)
# there doesn't seem to be any difference between the .csv and .txt files

table(stanford_diagnoses$Sample_Type)

# check order is same
colnames(mcols(stanford_input_Gr)) == paste('X', stanford_diagnoses$X, sep = '')


colnames(mcols(stanford_input_Gr))

colnames(mcols(stanford_input_Gr)) %in% paste('X', stanford_diagnoses$X, sep = '')


to_keep <- colnames(mcols(stanford_input_Gr)) %in% paste('X', stanford_diagnoses$X, sep = '')
mcols(stanford_input_Gr) <- mcols(stanford_input_Gr)[, to_keep]

# now is in right order, with just samples in diagnoses
colnames(mcols(stanford_input_Gr)) == paste('X', stanford_diagnoses$X, sep = '')

stanford_diagnoses <- stanford_diagnoses[, c('X', 'Sample_Type')]

# now keep only straight-forward diagnoses
valid_diagnoses <- c('Oncocytoma', 'RCC-Chromophobe', 'Normal kidney parenchyma', 'Clear cell RCC')
to_keep <- stanford_diagnoses$Sample_Type %in% valid_diagnoses

stanford_diagnoses <- stanford_diagnoses[to_keep, ]
mcols(stanford_input_Gr) <- mcols(stanford_input_Gr)[, to_keep]

dim(stanford_diagnoses)
dim(mcols(stanford_input_Gr))
# and check same order:
mean(paste('X', stanford_diagnoses$X, sep = '') == colnames(mcols(stanford_input_Gr)))

```


Now to put stanford_input_Gr in a dataframe and save them both
```{r}
to_save <- data.frame(stanford_input_Gr)
to_save <- to_save[, !(colnames(to_save) %in% c('end', 'width', 'strand'))] # remove unneeded columns
to_save$seqnames <- str_remove_all(as.character(to_save$seqnames), 'chr')

# check samples and features still in right order:
mean(paste0('X', stanford_diagnoses$X) == colnames(to_save)[3:length(colnames(to_save))])
mean(to_save$seqnames == seqnames_and_start$seqnames)
mean(to_save$start == seqnames_and_start$start)

write.csv(to_save, 'Brennandata/preprocessed_validation_beta.csv')
write.csv(stanford_diagnoses, 'Brennandata/preprocessed_validation_diagnoses.csv')



```


Now plot PCA and check looks ok
```{r}
# looking at the PCA to see if there are the normal clusters
no_nas <- na.omit(to_save[, 3:ncol(to_save)])

pca_res <- prcomp(t(no_nas))

pca_data <- as.data.frame(pca_res$x)
library(ggplot2)
ggplot(data = pca_data) +
  geom_point(aes(x = PC1, y = PC2, colour = stanford_diagnoses$Sample_Type)) 
# looks ok, have some clusters

```


Now I will run them through validate validation data, and create UMAPs of them with train/test data, to look for potential issues.


# Now for pre-processing of GSE61441 KIRC external data.

First read in the data:
```{r}
library(data.table)
GSE_data <- fread('Wei_external/GSE61441_Unmethylated_and_methylated_signal_intensities.csv', data.table = FALSE)

# Need to calculate the beta values from the raw count data
library(stringr)
met_cols <- str_detect(colnames(GSE_data), ' Methylated')
unmet_cols <- str_detect(colnames(GSE_data), ' Unmethylated')
met_data <- GSE_data[, met_cols ]
unmet_data <- GSE_data[, unmet_cols ]

head(met_data)
head(unmet_data)

# the sample orders are still the same:
mean(str_remove(colnames(met_data), ' Methylated signal') == str_remove(colnames(unmet_data), ' Unmethylated Signal'))

# so we can do the beta = met/(met + unmet) calculation
betas <- met_data/(met_data + unmet_data)
colnames(betas) <- str_remove(colnames(betas), ' Methylated signal')
head(betas)

# manual check:
i = 4
j = 3
betas[i, j] == met_data[i, j]/(met_data[i, j] + unmet_data[i, j])

```

Now to turn the probe names into probe locations
```{r}
# need to get probe positions from a harmonised 450k annotation
probe_location_data <- fread('jhu-usc.edu_BLCA.HumanMethylation450.24.lvl-3.TCGA-XF-AAMW-01A-11D-A42F-05.gdc_hg38.txt', data.table = F)
probe_location_data <- probe_location_data[, c('Composite Element REF', 'Chromosome', 'Start', 'End')]

betas$probe = GSE_data$ID_REF
beta_data_probe_locs <- merge(probe_location_data, betas, by.x = 'Composite Element REF', by.y = 'probe', all = TRUE, sort = FALSE)

beta_GRanges <- GRanges(seqnames = beta_data_probe_locs$Chromosome, strand ='*', ranges = IRanges(start = beta_data_probe_locs$Start, end = beta_data_probe_locs$End))
mcols(beta_GRanges) <- beta_data_probe_locs[, 5:ncol(beta_data_probe_locs)] # all the samples


# we saved the positions of the processed training data. lets read them in here
seqnames_and_start <- fread('training_testing_maxgap_50_newdata_processed_m_vals.csv', select = c('seqnames', 'start'), sep = ',')

input_features_Gr <- GRanges(seqnames = seqnames_and_start$seqnames, ranges = IRanges(start = seqnames_and_start$start, end = seqnames_and_start$start), strand = '*')
input_features_Gr
seqlevelsStyle(input_features_Gr) <- seqlevelsStyle(beta_GRanges)

# now to find overlaps
overlaps <- findOverlaps(beta_GRanges, input_features_Gr)
overlaps

mean(1:length(input_features_Gr) %in% subjectHits(overlaps)) # good so we overlap for every input_feature
subjectHits(overlaps)[duplicated(subjectHits(overlaps))] # but multiple chopera positions can map to one input_feature position
# lets just delete one, it should be very similar methylation as it is the same position

# need to order because for some reason becomes unordered
ordered_overlaps <- arrange(as.data.frame(overlaps), subjectHits)
ordered_overlaps

beta_input_Gr <- beta_GRanges[ordered_overlaps$queryHits][!duplicated(granges(beta_GRanges[ordered_overlaps$queryHits]))]

# check features are correct
mean(seqnames(beta_input_Gr) == paste0('chr', seqnames_and_start$seqnames))
mean(start(beta_input_Gr) == seqnames_and_start$start) # yay both are 1

beta_input_Gr[1:5,1:5]

for (c in colnames(mcols(beta_input_Gr))) { # not too many NAs which is good
  print(sum(is.na(mcols(beta_input_Gr)[, c])))
}



```

Now get the diagnoses and save 
```{r}
tumour_or_normal <- str_sub(colnames(mcols(beta_input_Gr)), 12, 13)
diagnoses <- data.frame(sample_id = colnames(mcols(beta_input_Gr)), label = ifelse(tumour_or_normal == 'T', 'KIRC', 'Normal'))


to_save <- data.frame(beta_input_Gr)
to_save <- to_save[, !(colnames(to_save) %in% c('end', 'width', 'strand'))] # remove unneeded columns
to_save$seqnames <- str_remove_all(as.character(to_save$seqnames), 'chr')

# chekc samples and features still in right order:
mean(paste0('X', diagnoses$sample_id) == colnames(to_save)[3:length(colnames(to_save))])
mean(to_save$seqnames == seqnames_and_start$seqnames)
mean(to_save$start == seqnames_and_start$start)

write.csv(to_save, 'Wei_external/preprocessed_validation_beta.csv')
write.csv(diagnoses, 'Wei_external/preprocessed_validation_diagnoses.csv')


```

Now plot PCA and check looks ok
```{r}
# looking at the PCA to see if there are the normal clusters
no_nas <- na.omit(to_save[, 3:ncol(to_save)])

pca_res <- prcomp(t(no_nas))

pca_data <- as.data.frame(pca_res$x)
library(ggplot2)
ggplot(data = pca_data) +
  geom_point(aes(x = PC1, y = PC2, colour = diagnoses$label)) 
# looks good, have 2 clusters ish

```


Now I will run them through validate validation data, and create UMAPs of them with train/test data, to look for potential issues.



# Pre-processing of Evelonn data - normalised beta values

```{r}
evelonn_data <- readRDS('Evelonn_external/evelonndataset_450k_downloadedGEO_210825.rds')
evelonn_data

evelonn_sample_details <- read.table('Evelonn_external/evelonn_sample_details.csv', sep=',', header=T)

colnames(evelonn_data) <- str_remove(colnames(evelonn_data), 'GSE113501_series_matrix.txt.gz.')

evelonn_data

wanted_probes_and_positions <- read.table('training_testing_maxgap_50_newdata_processed_probe_names.csv', sep=',', header=T)


nas <- (is.na(evelonn_data))
sum(nas)
mean(nas) # a lot more NAs than the raw data (processed above)
library(pheatmap)
nas_subset <- nas[, 1:5000]
pheatmap(as.matrix(nas_subset)*1, cluster_rows = F, cluster_cols=F) # NAs are different for multiregion samples



# so now we can select these probes
evelonn_data_wanted_probes <- evelonn_data[, as.character(wanted_probes_and_positions$Composite.Element.REF)]

evelonn_data_wanted_probes

sum(colMeans(is.na(evelonn_data_wanted_probes)) == 1)

mean(colnames(evelonn_data_wanted_probes) == as.character(wanted_probes_and_positions$Composite.Element.REF))


# transpose
evelonn_data_wanted_probes_t <- t(evelonn_data_wanted_probes)
evelonn_data_wanted_probes_t <- data.frame(evelonn_data_wanted_probes_t)


# now save
evelonn_sample_details$diagnosis <- ifelse(str_detect(evelonn_sample_details$stage, 'TumorFree'), 'Normal', 'ccRCC')
write.csv(evelonn_data_wanted_probes_t, 'Evelonn_external/preprocessed_validation_beta_from_normalised_data.csv')
write.csv(evelonn_sample_details, 'Evelonn_external/preprocessed_validation_beta_diagnoses_from_normalised_data.csv')
```


# Now check this data looks right
```{r fig.width=10}
evelonn_processed <- fread('Evelonn_external/preprocessed_validation_beta_from_normalised_data.csv')
evelonn_diagnoses <- fread('Evelonn_external/preprocessed_validation_beta_diagnoses_from_normalised_data.csv', header = TRUE)

evelonn_processed[1:5,1:10]


# comparing the probes to some training probes to see if they look vaguely similar
train <- data.frame(fread('training_testing_maxgap_50_newdata_processed_m_vals.csv'))
row.names(train) <- paste(train$seqnames, train$start)
train <- train[, 4:ncol(train)]
train_diag <- data.table(fread('training_testing_diagnoses_maxgap_50_newdata.csv'))
train_diag <- train_diag[3:nrow(train_diag), ]
train_diag$overallpath[500:507] 
train[1:10, 500:507] 

betas <- evelonn_processed[1:10, 20:29]

log2(betas/(1-betas)) # these look much more similar than the previous non-normalised processed data


# looking at normals:
betas <- evelonn_processed[1:10, 116:120]
log2(betas/(1-betas)) 
train_normals <- train[, train_diag$overallpath == 0]
# yes these look good too

# looking at the PCA to see if there are the normal clusters
no_nas <- na.omit(evelonn_processed[, 2:ncol(evelonn_processed)])

pca_res <- prcomp(t(no_nas))

pca_data <- as.data.frame(pca_res$x)
evelonn_diagnoses <- as.data.frame(evelonn_diagnoses)
row.names(evelonn_diagnoses) <- evelonn_diagnoses$accession
library(ggplot2)
library(ggrepel)
ggplot(data = pca_data) +
  geom_point(aes(x = PC1, y = PC2, colour = evelonn_diagnoses$diagnosis, shape=str_detect(evelonn_diagnoses$name, 'Pat')), size=3)


incorrect_evelonns <- read.csv('../figs_xgboost_validation/incorrectly_predicted_Evelonn_samples_model_3.csv')

# now look at where the incorrectly predicted samples are
ggplot(data = pca_data) +
  geom_point(aes(x = PC1, y = PC2, shape=str_detect(evelonn_diagnoses$name, 'Pat'), color=evelonn_diagnoses$accession %in% incorrect_evelonns$colname), size = 3) +
  theme(legend.title = element_text( size=8), legend.text=element_text(size=8))


# get more details about the samples, like metastasis status
gset <- getGEO("GSE113501", GSEMatrix =TRUE, getGPL=FALSE) # WARNING: takes a while!!
if (length(gset) > 1) idx <- grep("GPL13534", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]
sample_details <- data.frame(accession=gset$geo_accession, stage=gset$`tnm stage:ch1`, name=gset$title, metastasis_status=gset$`progress status (m0-pf:ch1`, gender=gset$characteristics_ch1.1, age=gset$`age:ch1`)
mean(sample_details$accession == evelonn_diagnoses$accession) # check same order


# metastasis status:
ggplot(data = pca_data) +
  geom_point(aes(x = PC1, y = PC2, colour = str_sub(sample_details$metastasis_status, 164), shape=str_detect(evelonn_diagnoses$name,'Pat')), size=3) # doesn't cluster with metastasis stage...

# stage:
ggplot(data = pca_data) +
  geom_point(aes(x = PC1, y = PC2, colour = sample_details$stage, shape=str_detect(evelonn_diagnoses$name,'Pat')), size=3) 

# gender:
ggplot(data = pca_data) +
  geom_point(aes(x = PC1, y = PC2, colour = sample_details$gender, shape=str_detect(evelonn_diagnoses$name,'Pat')), size=3) 

# age:
ggplot(data = pca_data) +
  geom_point(aes(x = PC1, y = PC2, colour = as.integer(as.character(sample_details$age)), shape=str_detect(evelonn_diagnoses$name,'Pat')), size=3)



# working theory: the right cluster are different somehow, maybe mistaken chRCC or metastases that our classifier cannot recognise. I predict the 3 single ccRCC samples will also be misclassified
# this is indeed correct

right_cluster_gsms <- row.names(pca_data[pca_data$PC1 > 20, ])

evelonn_sample_details[evelonn_sample_details$accession %in% right_cluster_gsms, ]
```

# Does the outlying cluster associate with purity?
```{r fig.width=10}
purity_ests <- readRDS('Evelonn_external/purity_evelonn_forizzy_using156normalisedprobes_210914.rds')
write.csv(purity_ests, 'Evelonn_external/purity_evelonn_forizzy_using156normalisedprobes_210914.csv')
evelonn_diagnoses
evelonn_diagnoses_with_purity <- left_join(evelonn_diagnoses, purity_ests, by = c('accession'='sample.id'))
mean(evelonn_diagnoses$accession == evelonn_diagnoses_with_purity$accession)

ggplot(data = pca_data) +
  geom_point(aes(x = PC1, y = PC2, colour = evelonn_diagnoses_with_purity$purity, shape=str_detect(evelonn_diagnoses$name,'Pat')), size=3)

```




# Code to download Evelonn data from GEO:
```{r}
library(GEOquery)
library(limma)

# load series and platform data from GEO
  
gset <- getGEO("GSE113501", GSEMatrix =TRUE, getGPL=FALSE)
if (length(gset) > 1) idx <- grep("GPL13534", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]

gset

test <- data.frame(gset)

test

# View(phenoData(gset)@data)

sample_details <- data.frame(accession=gset$geo_accession, stage=gset$`tnm stage:ch1`, name=gset$title, metastasis_status=gset$`progress status (m0-pf:ch1`)

write.table(sample_details, 'Evelonn_external/evelonn_sample_details.csv', sep=',', row.names=F)


```



